\textls[-0]{\thetitle}

### 2023-03-22 canonicitymodel.tex
comments

\begin{align*}
  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ (also sets indexed by closed substitution)}\\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{j}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} \\
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\delta}{\Delta}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)} \denotesC{\goodCtx{\Dl}{}}(\delta \circ \gm) \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
  \denotesC{\goodType{\Gm}{\sub{T}{\sigma}}{}}(\gm)(\gm')(t) &= \denotesC{T}(\sigma \circ \gm)(\denotes{\sigma}(\gm)(\gm'))(t)\\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= \{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \ | \text{ when } t \equiv \false  \} \quad \text{disjoint union} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \\
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')(\fst{t})} \denotesC{B}(\SubstExt{\gm}{\fst{t}})((\gm', u'))(\snd{t}) \\
  \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \{ (\gamma' , t') : \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  \} \\
  \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm)(\gm') &= \star\\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm)(\gm') &= \star^1 \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm)(\gm') &= \star^2 \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) ((\gm' , u')) \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}}(\gm)(\gm') &= \denotesC{t}(\pi_1 \gm)(\gm'[0])(\pi_2 \gm)(\gm'[1]) \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}}(\gm)(\gm') &= (\denotesC{a}(\gm)(\gm') , \denotesC{b}(\gm)(\gm') )\\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[0] \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[1]\\
  \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\}  \to \Set{j} \\
  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm)(\gm') &= \denotesC{T}(\gm)(\gm') \\ 
  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm)(\gm')(t) &= \denotesC{T}(\gm)(\gm')(t) \\
\end{align*}

@@

%The difficult part of the proof-relevant logical-relations model is the
%treatment of propositions---
Proof-relevant model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}
a proposition is encoded as a subsingleton data. For
example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
| \text{ when } t \equiv \false  \}$ represents a disjoint union of two
subsingleton sets, where the first one is either a singleton if we have a proof of
$t \equiv \true $ or an empty set otherwise.
In other words,
$\star^1$ witnesses a proof that $t \equiv \true$.
Here, $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.

@@

Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$
that returns the syntax piece $S$ as well.\footnote{This is also called
glued interpretation in \citet{sterling2019algebraic}.}
Thus, we can have a
function $\Glued{\Pi}$ such that $\GluedPi{\denotesCC{A}}{\denotesCC{B}} =
\denotesCC{\TyPi{A}{B}}$ now that the syntax is available.
Similarly, there are functions $\Glued{\Sigma}$, $\Gpair{a}{b}$  and
$\GSubstExt{\gm}{t}$ for dependent pair types, dependent pairs, and
substitution extension (and more for other constructions).
Furthermore, given $\Glued{S}$ (i.e., the syntax and its semantic interpretation),
we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.\YZ{why not just $\denotesC{S}$?}\EDJreply{Using superscript C is shorter and more precise -- we are not "reconstructing the C model from the syntax, but taken out the C model directly from the pair". Look at the canonicity model where superscript c is used, if change to $\denotesC{-}$ on them, it is not really well-founded recursion, technically speaking (only equivalent to one).}
% These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT. What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.

>>>>

  \denotesC{\goodCtx{\Gm}{k}} \text{ is a function }&: \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ (also sets indexed by closed substitution)}\\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{j}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}}\prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)}\{t : \goodTerm{\cdot}{t}{\sub{T}{\gm}}\}\to\Set{j} \\
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\delta}{\Delta}} \text{ is a dependent function }&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}(\gm)} \denotesC{\goodCtx{\Dl}{}}(\delta \circ \gm) \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} \text{ is a dependent function}&: \prod_{\goodSub{\cdot}{\gm}{\Gm}} \prod_{\gm' \in \denotesC{\goodCtx{\Gm}{}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)(\gm')(\sub{t}{\gm})\\
  \denotesC{\goodType{\Gm}{\sub{T}{\sigma}}{}}(\gm)(\gm')(t) &= \denotesC{T}(\sigma \circ \gm)(\denotes{\sigma}(\gm)(\gm'))(t)\\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(\gm')(t) &= \{\star \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm)(\gm')(t) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm)(\gm')(t) &= 
    \begin{cases}
      \{\star^1\} & \text{if } t \equiv \true \\
      \{\star^2\} & \text{if } t \equiv \false
    \end{cases}\YZ{Is this more accurate than 'disjoint union'?}
%   \setc{\star^1}{ \text{ when } t \equiv \true } \dot{\cup} \setc{\star^2}{\text{ when } t \equiv \false } \quad \text{disjoint union} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm)(\gm')(t) &= \{\star \ |  \text{ when } t \equiv \eqrefl{\sub{a}{\gm}} \text{ and } \sub{a}{\gm} \equiv \sub{b}{\gm} \} \YZ{What if the conditions do not hold? Empty set?}\\
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}}(\gm)(\gm')(t) &= \prod_{\goodTerm{\cdot}{u}{\sub{A}{\gm}}} \ \ \prod_{u' \in \denotesC{A}(\gm)(\gm')(u)} \denotesC{B}(\SubstExt{\gm}{u})((\gm', u'))(\sub{\app{t}}{\SubstExt{"id"}{u}})\\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}}(\gm)(\gm')(t) &= \sum_{u' \in \denotesC{A}(\gm)(\gm')(\fst{t})} \denotesC{B}(\SubstExt{\gm}{\fst{t}})((\gm', u'))(\snd{t}) \\
  \denotesC{\goodCtx{\cdot}{}}(\gm) &= \{\star\} \quad \text{  a singleton set } \\ 
  \denotesC{\goodCtx{\Gm, T}{}}(\gm_t) &= \setc{ (\gamma' , t') }{ \gm' \in \denotesC{\goodCtx{\Gm}{}}(\pi_1 \gm_t), t' \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}(\pi_1 \gm_t)(\gm')(\pi_2 \gm_t)  } \\
  \denotesC{\goodTerm{\Gm}{\sub{t}{\sigma}}{\sub{T}{\sigma}}}(\gm)(\gm') &= \denotesC{t}(\sigma \circ \gm)(\denotesC{\sigma}(\gm)(\gm')) \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm)(\gm') &= \star\\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm)(\gm') &= \star^1 \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm)(\gm') &= \star^2 \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{a}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^1 \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm)(\gm') &= \denotesC{b}(\gm)(\gm') \quad \text{ if } \denotes{c}(\gm)(\gm') = \star^2 \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm)(\gm') &= \star\\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}}(\gm)(\gm') &= \lambda u \lambda u' . \denotesC{t}(\SubstExt{\gm}{u}) ((\gm' , u')) \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}}(\gm)(\gm') &= \denotesC{t}(\pi_1 \gm)(\gm'[0])(\pi_2 \gm)(\gm'[1]) \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}}(\gm)(\gm') &= (\denotesC{a}(\gm)(\gm') , \denotesC{b}(\gm)(\gm') )\\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[0] \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}}(\gm)(\gm') &= \denotesC{t}(\gm)(\gm')[1]\\
  \denotesC{\goodType{\Gm}{\cU_j}{j+1}}(\gm)(\gm')(T) &= \setc{t} { \goodTerm{\cdot}{t}{\El{T}} }  \to \Set{j} \\
  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm)(\gm') &= \denotesC{T}(\gm)(\gm') \\ 
  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm)(\gm')(t) &= \denotesC{T}(\gm)(\gm')(t) \\

@@

The difficult part of the proof-relevant logical-relations model is the
treatment of propositions.
The model treats proposition in a special way : \EDJ{Without this first sentence, this paragraph seems unmotivated.}\YZreply{Why is it 'difficult'? Is there a more appropriate adjective than 'difficult'?}
a proposition is encoded as a sub\-singleton. For
example, $\{\star^1 \ | \text{ when } t \equiv \true  \} \dot{\cup} \{\star^2 \
| \text{ when } t \equiv \false  \}$ represents a disjoint union of two
sub\-singleton sets, where the first one is either a singleton if we have a proof of
$t \equiv \true $ or an empty set otherwise.
In other words,
$\star^1$ witnesses a proof that $t \equiv \true$.
Here, $\star$, $\star^1$, $\star^2$ are just some arbitrary fixed elements.

@@

Thus, we define a new denotation $\denotesCC{S} \coloneq (S, \denotesC{S})$
that also returns the syntax piece $S$.\footnote{This is also called
glued interpretation in \citet{sterling2019algebraic}.}
Thus, we can have a
function $\Glued{\Pi}$ such that $\GluedPi{\denotesCC{A}}{\denotesCC{B}} =
\denotesCC{\TyPi{A}{B}}$ now that the syntax is available.
Similarly, there are functions $\Glued{\Sigma}$, $\Gpair{a}{b}$, and
$\GSubstExt{\gm}{t}$ for dependent pair types, dependent pairs, and
substitution extension (and more for other constructions).
Furthermore, given $\Glued{S}$ (i.e., the syntax and its semantic interpretation),
we use $\Gmodel{\Glued{S}}$ to mean the latter of the two.%\YZ{why not just $\denotesC{S}$?}\EDJreply{Using superscript C is shorter and more precise -- we are not "reconstructing the C model from the syntax, but taken out the C model directly from the pair". Look at the canonicity model where superscript c is used, if change to $\denotesC{-}$ on them, it is not really well-founded recursion, technically speaking (only equivalent to one).}
% These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT. What's more, during the construction of the model for \TT, we will use $\Glued{(\cdot)}$ when we can.
